---
title: "Online Estimation of Surrogate Measures and its Application to Transportation Safety Analysis"
author: "Yilin Zhang"
date: "2021/3/23"
output:
  prettydoc::html_pretty:
    theme: architect 
    highlight: github
---


## Background

### Data Source:

Traffic crashes are rare events, with an average rate of one crash 6.8 per million vehicle miles traveled in the US. The **naturalistic driving study (NDS)** provides an unprecedented opportunity to  evaluate crash risk. NDSs are characterized by  continuously recording driving information, such as three-dimensional Inertial Measurement Unit (IMU) acceleration, GPS and multi-channel video recordings[1].  

- Accurately identifying crashes with robustness
- Benefitting further understanding of driving behavior 
- With the overall goal of reducing accidents 

This experiment  uses the **Second Strategic Highway Research Program (SHRP 2) NDS**, the largest NDS to-date, with more than 3,400 participants and 1 million hours of continuous driving data[2].



### Surrogate Measures:

To mitigate the rarity and insufficiency for traffic crash \citep{guo2010near}, a broad spectrum of studies investigate identification and development of various surrogate measures, which helps to supplement the analytical data and enhance the reliability of safety analysis.

To be useful for transportation safety applications, a **surrogate measure** should[4] 

- related in a predictable and reliable way to crashes
- converting the non-crash events into a corresponding crash frequency and/or severity
- has physical meaning

Much of the empirical work is framed by detecting the jerky driving behaviors based on the **acceleration information**. The premise is the positive correlation between **elevated gravitational-force events and crash propensity**. While such an attempt is sensitive to the unstable HGF threshold and can easily generate too many false positives.
 
### Aim of Our Work

Our core concern is to propose surrogates based on three-dimension of acceleration.

- Propose new surrogate measures with large power in detecting crash
- Implement the surrogate measures with online version
 
## Exploratory Data Analysis

We first import data and combine them together.

- Select 10 segments, with 5 crash ones and 5 non-crash ones. 
- Each segments have around 300 time points.

```{r message=FALSE, warning=FALSE}
library(e1071)
library(plyr)
library(dplyr)
library(ggplot2)
library(reshape2)
library(gridExtra)
```


```{r}
dir_crash <- "C:/Users/yzhang/Desktop/appliedstat/OnlineSurrogate/data/segment/crash/"
setwd(dir_crash)
f_crash <- list.files(dir_crash)
data_crash <- ldply(f_crash[1:6], read.csv, header=TRUE)
data_crash <- mutate(data_crash, crash = 1)


dir_base <- "C:/Users/yzhang/Desktop/appliedstat/OnlineSurrogate/data/segment/base/"
setwd(dir_base)
f_base <- list.files(dir_base)
data_base <- ldply(f_base[1:6], read.csv, header=TRUE)
data_base <- mutate(data_base, crash = 0)

dataset <- bind_rows(data_base, data_crash)
```


### Basic description for driving segments


```{r}
temp <- data_base %>% select(file_id, X, accel_x, accel_y, accel_z)
temp <- melt(temp, id.vars = c("file_id", "X"))


temp %>%
  ggplot(aes(x=X, y=value, group=variable, color=variable)) +
  geom_line(size=0.7) +
  facet_wrap("file_id")+
  ggtitle("Acceleration for baseline segments") + theme_minimal()+
  theme(plot.title = element_text(size = 15, hjust = 0.5),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())

```


```{r message=FALSE, warning=FALSE}
temp <- data_crash %>% select(file_id, X, accel_x, accel_y, accel_z)
temp <- melt(temp, id.vars = c("file_id", "X"))

temp %>%
  ggplot(aes(x=X, y=value, group=variable, color=variable)) +
    geom_line(size=0.7) +
    facet_wrap("file_id")+
    ggtitle("Acceleration for crash segments") + theme_minimal()+
    theme(plot.title = element_text(size = 15, hjust = 0.5),
          axis.title.x=element_blank(),
          axis.title.y=element_blank())
```


Here, we give focus on some crash samples.

```{r message=FALSE, warning=FALSE}
temp <- data_crash %>% 
  filter(file_id == "63772054") %>%
  select(X, accel_x, accel_y, accel_z)
temp <- melt(temp, id.vars = c("X"))


temp %>%
  ggplot(aes(x=X, y=value, group=variable, color=variable)) +
  geom_line(size=1.0) +
  ggtitle("Acceleration for crash segment 63772054") + theme_minimal() +
  theme(plot.title = element_text(size = 15, hjust = 0.5),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

Mild crash:

- Slam on the brakes and slow down to stop   
 

```{r message=FALSE, warning=FALSE}
temp <- data_crash %>% 
  filter(file_id == "37159529") %>%
  select(X, accel_x, accel_y, accel_z)
temp <- melt(temp, id.vars = c("X"))


temp %>%
  ggplot(aes(x=X, y=value, group=variable, color=variable)) +
  geom_line(size=1.0) +
  ggtitle("Acceleration for crash segment 37159529") + theme_minimal() +
  theme(plot.title = element_text(size = 15, hjust = 0.5),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

Heavy crash:

- Slam on the brakes 
- No stop and straight collision
- Hit the steering wheel

### Crash V.S. Non-Crash


```{r}
f_dir <- "C:/Users/yzhang/Desktop/appliedstat/online/select/crash"
setwd(f_dir)
data_crash <- ldply(list.files(f_dir), read.csv, header=TRUE)
data_crash <- mutate(data_crash, crash = 1)

f_dir <- "C:/Users/yzhang/Desktop/appliedstat/online/select/baseline/"
setwd(f_dir)
data_non <- ldply(list.files(f_dir), read.csv, header=TRUE)
data_non <- mutate(data_non, crash = 0)

dataset <- bind_rows(data_non, data_crash)
str(dataset)
```


We plot the density for crash and non-crash data in three dimension and find

- the rare event account for a small part but with extreme value
- hard to distinguish crash and non-crash only from extreme value

```{r message=FALSE}
temp <- dataset %>% 
  select(crash, accel_x, accel_y, accel_z)
log_d1 <- function(x){
  return(log(abs(x)+1)*sign(x))
}
log_d2 <- function(x){
  return(log(abs(x+1)+1)*sign(x)-1)
}
temp["accel_x"] <- lapply(temp["accel_x"], log_d1)
temp["accel_y"] <- lapply(temp["accel_y"], log_d1)
temp["accel_z"] <- lapply(temp["accel_z"], log_d2)
temp <- melt(temp, id.vars = c("crash"))
temp$crash <- factor(temp$crash)
ggplot(temp, aes(x=variable, y=value, fill=crash)) +
  geom_violin()+
  ggtitle("Acceleration density in three-dimension") + theme_minimal()+
  theme(plot.title = element_text(size = 15, hjust = 0.5))
```


We consider a new surrogate, **standard deviation of acceleration**.
```{r message=FALSE}
temp <- dataset %>% 
  select(crash, file_id, accel_x, accel_y, accel_z) %>%
  group_by(crash, file_id) %>%
  summarise(
    ac_std_x = sd(accel_x),
    ac_std_y = sd(accel_y),
    ac_std_z = sd(accel_z)
  )%>% 
  select(crash, ac_std_x, ac_std_y, ac_std_z) %>%
  melt(id.vars = c("crash"))
temp$crash <- factor(temp$crash)
p1 <- ggplot(temp, aes(x=variable, y=value, fill=crash)) +
  geom_violin()+
  ggtitle("Standard deviation") +     
  theme_minimal()+
  theme(plot.title = element_text(size = 15, hjust = 0.5))
```


And a new surrogate, **skewness of acceleration**.
```{r message=FALSE}
temp <- dataset %>% 
  select(crash, file_id, accel_x, accel_y, accel_z) %>%
  group_by(crash, file_id) %>%
  summarise(
    ac_ske_x = skewness(accel_x, type = 1),
    ac_ske_y = skewness(accel_y, type = 1),
    ac_ske_z = skewness(accel_z, type = 1)
  )%>% 
  select(crash, ac_ske_x, ac_ske_y, ac_ske_z) %>%
  melt(id.vars = c("crash"))
temp$crash <- factor(temp$crash)
p2 <- ggplot(temp, aes(x=variable, y=value, fill=crash)) +
  geom_violin() +
  ggtitle("Skewness") + 
  theme_minimal()+
  theme(plot.title = element_text(size = 15, hjust = 0.5))
```

And a new surrogate, **skewness of acceleration**.
```{r message=FALSE}
temp <- dataset %>% 
  select(crash, file_id, accel_x, accel_y, accel_z) %>%
  group_by(crash, file_id) %>%
  summarise(
    ac_kur_x = kurtosis(accel_x, type = 1),
    ac_kur_y = kurtosis(accel_y, type = 1),
    ac_kur_z = kurtosis(accel_z, type = 1)
  )%>% 
  select(crash, ac_kur_x, ac_kur_y, ac_kur_z) %>%
  melt(id.vars = c("crash"))
temp$crash <- factor(temp$crash)
p2 <- ggplot(temp, aes(x=variable, y=value, fill=crash)) +
  geom_violin() +
  ggtitle("Skewness") + 
  theme_minimal()+
  theme(plot.title = element_text(size = 15, hjust = 0.5))
```

```{r}
l <- list(p1, p2)
grid.arrange(grobs = l, ncol = 2)
```


From the figures above, standard deviation or skewness can be better surrogates.

## Future Work on Model

Our core concern is to propose surrogates based on three-dimension of acceleration.

### New surrogate measures 
We are going to propose new surrogate measures that

- has large power in detecting crash 
- has low false positives

The potential surrogates may be

- standard deviation of acceleration
- skewness of acceleration
- kurtosis of acceleration

We use a simple model "Single Index model", to convert the surrogates to the probability of crash.
$$f(t) = g(\beta^{T}t)+\epsilon$$

### Implement surrogates with online version

Considering that all the surrogates we proposed, can be estimated using U statistics or function of U statistics. Thus, we derive a general frame work for online U statistics.


- In the first step, we calculate the U statistics based on all initial data, which is 
$$\hat\theta^{(0)} = (C_{n}^{m})^{-1}\sum_{ \{i_1,\ldots,i_m\}\in I_0} h(X_{i_1},\ldots,h_{i_m}).$$

- To refresh the U statistics in time $t$, we use the following iterative formula. 
$$R^{(t)} = \sum\limits_{\{i_1,\ldots,i_m\}\in I_t}h(X_{i_1},\ldots,X_{i_m}) , $$
$$C^{(t)} = \sum_{k=1}^{m-1}\sum\limits_{\{i_1,\ldots,i_k\}\in J_{t-1}, \{i_{k+1},\ldots,i_m\}\in I_{t}}h(X_{i_1},\ldots,X_{i_m})$$    

$$\hat\theta^{(t)} = \left\{(t+1)C_n^m-tC_r^m\right\}^{-1}\left[\left\{tC_n^m-(t-1)C_r^m\right\}\hat\theta^{(t-1)} + C^{(t)} + R^{(t)}\right]$$
                                            
- By transforming $\hat\theta^{(t)}$ as $T(\hat\theta^{(t)})$, we derive the new estimates for $T(\theta)$.







## Reference

[1] Feng Guo. Statistical methods for naturalistic driving studies. Annual review of statisticsand its application, 6:309–328, 2019.

[2]  Thomas  A  Dingus,  Feng  Guo,  Suzie  Lee,  Jonathan  F  Antin,  Miguel  Perez,  MindyBuchanan-King, and Jonathan Hankey. Driver crash risk factors and prevalence evalu-ation using naturalistic driving data. Proceedings of the National Academy of Sciences,113(10):2636–2641, 2016.

[3] Feng Guo, Sheila G Klauer, Jonathan M Hankey, and Thomas A Dingus.  Near crashesas  crash  surrogate  for  naturalistic  driving  studies. Transportation  Research  Record,2147(1):66–74, 2010.

[4]  Andrew P Tarko.  Surrogate measures of safety.  InSafe Mobility:  Challenges, Method-ology and Solutions. Emerald Publishing Limited, 2018.



